---
title: "Untitled"
output: html_document
date: "2024-08-31"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##### Question-2

# 1)
```{r}
setwd("~/Downloads/M A S T E R S /Sem-3/Part-1/Visualization/labs/lab-1")

data = read.table("SENIC.txt",header = FALSE)

```

# 2)
```{r}

outlier <- function(variable){
  
  quants <- quantile(variable)
  
  Q1 <- quants[2]
  Q3 <- quants[4]
  
  upper <- Q3+1.5*(Q3-Q1)
  lower <- Q1-1.5*(Q3-Q1)
  
  value1 <- which(variable > upper)
  value2 <- which(variable < lower)
  
  final <- c(value1,value2)
  
  return(final)
}

```

# 3)
```{r}
library(ggplot2)

ind <- outlier(data$V4)
values <- data$V4[ind]

viz <- ggplot(data = data, aes(x = V4)) +  geom_density(kernel = "gaussian") + geom_point(data = data.frame(values), aes(x = values, y = 0),shape = 23, color = "red") + labs( x = "Infection risk", y = "Density")

viz 
```
We can see from the above graph, it resembles the shape of a normal distribution which majority of the observations around 4.5.There are only a few outliers i.e 5 outliers to be exact.  

# 4) 
```{r}
library(gridExtra)

col_names <- c("V2","V3","V5","V6","V7","V10","V11","V12")

col_real <- c("Length of Stay ", "Age", "Routine Culturing Ratio", "Routine Chest X-ray Ratio", "Number of Beds ", "Average Daily Census", "Number of Nurses", "Available Facilities & Services")

final<-list()

i <- 1
for(cols in col_names){
  
  ind <- outlier(data[[cols]])
  values <- data[[cols]][ind]
  
  out <- ggplot(data = data, aes(x = .data[[cols]])) +  geom_density(kernel = "gaussian") + geom_point(data = data.frame(values), aes(x = values, y = 0),shape = 23, color = "red") + labs( x = col_real[i] , y = "Density")
  
  final[[cols]]<-out
  
  i <- i + 1

}


plots <- arrangeGrob(grobs = final, ncol = 5, nrow = 2)  

grid.arrange(plots)


```
We notice outliers for all the variables except for the variable "Available Facilities & Services"(V12). We also see skewness in some variables like "Number of Beds" (V7), "Average Daily Census"(V10) and "Number of Nurses"(V11).

# 5) 
```{r}
ggplot(data,aes(V4,V11,color = factor(V7))) + geom_point() + labs( x = "Infection risk", y = "Number of Nurses")
```

We dont see anything prominent in the above graph to mention any insights. 

With regards to the color scale, having such a large color scale can be overwhelming plus having different shades of the colors makes it harder to analyze the graph to easily gain insights from it. 


# 6) 
```{r}
library(plotly)

ggplotly(viz)

```

Upon using plotly, we are able to hover over the graph and actually get the corresponding values for various locations on the graph like the outliers, we can hover over those outliers to easily identify its values. Additionally, we can also zoom in to a particular part of the graph we might be interested to analyse further. 


# 7) 
```{r}

ind <- outlier(data$V4)
values <- data$V4[ind]

quants <- quantile(data$V4)

Q1 <- quants[2]
Q3 <- quants[4]
  
upper <- Q3+1.5*(Q3-Q1)
lower <- Q1-1.5*(Q3-Q1)
  


data %>% plot_ly(x= ~V4, type = "histogram") %>% select(V4) %>% filter(V4 > upper | V4 < lower) %>% add_trace(type = "scatter", mode = "markers", marker = list(symbol = "diamond")) %>%  layout(xaxis = list(title = 'Infection risk'), showlegend = FALSE)

```



# 8) 
```{r}
library(shiny)


colnames(data) <- c("Identification Number","Length of Stay ", "Age","Infection Risk", "Routine Culturing Ratio", "Routine Chest X-ray Ratio", "Number of Beds ","Medical School Affiliation","Region", "Average Daily Census", "Number of Nurses", "Available Facilities & Services")

ui <- fluidPage(
  
   checkboxGroupInput("variable", "Variables to show:",c("Length of Stay ", "Age","Infection Risk", "Routine Culturing Ratio", "Routine Chest X-ray Ratio", "Number of Beds ", "Average Daily Census", "Number of Nurses", "Available Facilities & Services")),
   sliderInput("integer", "BW bandwidth:",min = 0, max = 100,value = 1), 
   
   plotOutput("plots")
   
)

# Define server actions
server <- function(input, output) {
  # Use input 
  # Save updates to output 
   
  output$plots <- renderPlot({
    
  final_2<-list()
  col_names_2 <- input$variable
  for(cols in col_names_2){
    
    ind <- outlier(data[[cols]])
    values <- data[[cols]][ind]
    
    out <- ggplot(data = data, aes(x = .data[[cols]])) +  geom_density(kernel = "gaussian",bw = input$integer) + geom_point(data = data.frame(values), aes(x = values, y = 0),shape = 23, color = "red") + labs( x = cols , y = "Density")
    
    final_2[[cols]]<-out
  
  }
  
  plots <- arrangeGrob(grobs = final_2, ncol = 5, nrow = 2)  

  #print(final_2)
  #print(input$variable)
  #grid.arrange(plots)
  plot(plots)
  #print(plots)
    
  })
  
}

# Run the application 
shinyApp(ui = ui, server = server)



```

The default bandwidth value that was provided is 1. And from the graphs, we can see that as we increase this value , the smoother the graphs becomes. 
